---
title: "Methods"
output: html_document
date: "2024-07-09"
---

# Pre proccesing
```{r, echo=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
library(tidymodels)
library(broom)

# Load the dataset
data <- read_csv('cleaned_smmh.csv')

# Filter out retired individuals because there are not many of them
data <- data %>% filter(occ_status != "Retired")

# Filter the data to include only Male and Female genders
data <- data %>% filter(gender %in% c("Male", "Female"))

# Convert avg_time_on_sm to numeric
data <- data %>%
  mutate(avg_time_on_sm = case_when(
    avg_time_on_sm == "Less than an Hour" ~ 0,
    avg_time_on_sm == "Between 1 and 2 hours" ~ 1,
    avg_time_on_sm == "Between 2 and 3 hours" ~ 2,
    avg_time_on_sm == "Between 3 and 4 hours" ~ 3,
    avg_time_on_sm == "Between 4 and 5 hours" ~ 4,
    avg_time_on_sm == "More than 5 hours" ~ 5
  ))

# Define custom labels for the subgraphs
illness_labels <- c(
  adhd_score = "ADHD Score",
  depression_score = "Depression Score",
  self_esteem_score = "Self-Esteem Score",
  anxiety_score = "Anxiety Score"
)

unique_values <- unique(data$gender)
print(unique_values)
```

# Dividing data into treatment and control gropus because the data is observational
```{r}
# Using Propensity Score Matching (PSM) to reduce selection bias when evaluating the time spent on social media on mental health scores. PSM is used for later inference on causation between the variables

# Create a binary indicator for social media usage (more than 2 hours)
data <- data %>%
  mutate(sm_usage = ifelse(avg_time_on_sm > 2, 1, 0))

# Split the data into training and testing sets
set.seed(123)
data_split <- initial_split(data, prop = 0.8, strata = sm_usage)
train_data <- training(data_split)
test_data <- testing(data_split)

# Convert sm_usage to factor in your dataset
train_data <- train_data %>%
  mutate(sm_usage = factor(sm_usage))

# Create a recipe for the propensity score model
ps_recipe <- recipe(sm_usage ~ overall_mental_health_score + occ_status, data = train_data) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_normalize(all_predictors())

# Specify the logistic regression model for propensity scores
ps_model <- logistic_reg() %>%
  set_engine("glm")

# Create a workflow for the propensity score model
ps_workflow <- workflow() %>%
  add_recipe(ps_recipe) %>%
  add_model(ps_model)

# Fit the propensity score model
ps_fit <- fit(ps_workflow, data = train_data)

# Predict propensity scores
train_data <- train_data %>%
  bind_cols(predict(ps_fit, new_data = train_data, type = "prob") %>% rename(propensity_score = .pred_1))

# Calculate the desired sample size for each group
sample_size <- min(table(train_data$sm_usage))

# Adjust the sampling size to min(table(sm_usage)) (the minimum between size of treatment group and size of control group)
train_data <- train_data %>%
  group_by(sm_usage) %>%
  sample_n(size = min(table(train_data$sm_usage))) %>%
  ungroup()
```

# Introduction on what are the possible relationships between different variables and mental health scores

##### Possible correlation between age and mental health
```{r}
# Create age groups in 5-year intervals from 15 to 60
data$age_group <- cut(data$age, breaks = seq(10, 60, by = 10), right = FALSE)

# Remove rows where age_group is NA
data <- data[!is.na(data$age_group), ]

# Create a box plot
ggplot(data, aes(x = age_group, y = overall_mental_health_score)) +
  geom_boxplot(fill = "#9B59B6", color = "black") +
    geom_jitter() +
  labs(x = "Age Group", y = "Overall Mental Health Score",
       title = "Distribution of Mental Health Scores by Age Group (5-Year Intervals)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels if needed
```
##### University students comapared to salaried workers and school students
```{r}
# Create a new column for count of people
temp_data <- data %>%
  group_by(occ_status, avg_time_on_sm) %>%
  summarise(count = n())

temp_data$count <- as.integer(temp_data$count)

plot <- ggplot(temp_data, aes(x = avg_time_on_sm, y = count, fill = occ_status)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = ifelse(count > 0, "", "")), position = position_stack(vjust = 0.5)) +  # Adjusted to show no labels
  facet_wrap(~ occ_status, scales = "fixed") +
  scale_fill_manual(values = custom_colors, name = "Occupation Status") +  # Use custom colors
    scale_x_continuous(breaks = 0:5) +  # Set x-axis breaks
  labs(title = "Count of People by Average Time on Social Media and Occupation Status",
       x = "Average Time on Social Media",
       y = "Count of People") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))

# Display the plot
print(plot)
```

```{r}
# Select specific columns for plotting
data_selected <- data %>%
  select(occ_status, adhd_score, depression_score, anxiety_score, self_esteem_score)

# Reshape data for plotting
data_long <- data_selected %>%
  pivot_longer(cols = ends_with("score"),
               names_to = "mental_health_condition", 
               values_to = "score")

# Plot grouped boxplot with facets
ggplot(data_long, aes(x = occ_status, y = score, fill = occ_status)) +
  geom_boxplot(outlier.shape = NA) +  # Remove outliers
  facet_wrap(~ mental_health_condition, scales = "free_y", ncol = 2, labeller = as_labeller(illness_labels)) +
  scale_fill_manual(values = custom_colors) +  # Use custom colors
  labs(title = "Distribution of Illness Scores by Occupation Status",
       x = "Occupation Status",
       y = "Score",
       fill = "Occupation Status") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    strip.text = element_text(size = 10, face = "bold")  # Customize facet strip text
  )
```

# Checking interaction effects between the different occupational groups
Main effects - Assuming relationship between average time on social media and mental health score does not vary by the occupational groups.
Interaction effect - Assuming relationship between average time on social media and mental health score varies by the occupational groups.
#### Main effects model
```{r}
# Fit model
main_effect_model <- lm(overall_mental_health_score ~ avg_time_on_sm + occ_status, data = data)

if (!is.factor(data$occ_status)) {
  data$occ_status <- factor(data$occ_status, levels = c("University Student", "School Student", "Salaried Worker"))
}

# Define levels explicitly for avg_time_on_sm and occ_status
avg_time_levels <- levels(data$avg_time_on_sm)
occ_status_levels <- levels(data$occ_status)

# Generate data for plotting
plot_data <- expand.grid(avg_time_on_sm = unique(data$avg_time_on_sm),
                         occ_status = levels(data$occ_status))

# Predict using the model
plot_data$predicted_score <- predict(main_effect_model, newdata = plot_data)

# Plotting with ggplot2
ggplot(plot_data, aes(x = factor(avg_time_on_sm), y = predicted_score, color = occ_status, group = occ_status)) +
  geom_line() +
  labs(title = "Main Effects",
    x = "Average Time on Social Media", y = "Predicted Mental Health Score") +
  scale_color_manual(values = c("blue", "green", "red"), name = "Occupation Status") +
  theme_minimal()
```
Same slope: Rate of change in mental health score as the average time of use increases does not vary between paintings by the occupational groups.
Different intercept: University students have consistently more score than the other occupational groups.
#### Interaction effect model
```{r}
# Fit your model (replace with your actual model fitting code)
intercation_effects_model <- lm(overall_mental_health_score ~ avg_time_on_sm * occ_status, data = data)

if (!is.factor(data$occ_status)) {
  data$occ_status <- factor(data$occ_status, levels = c("University Student", "School Student", "Salaried Worker"))
}

# Define levels explicitly for avg_time_on_sm and occ_status
avg_time_levels <- levels(data$avg_time_on_sm)
occ_status_levels <- levels(data$occ_status)

# Generate data for plotting
plot_data <- expand.grid(avg_time_on_sm = unique(data$avg_time_on_sm),
                         occ_status = levels(data$occ_status))

# Predict using the model
plot_data$predicted_score <- predict(intercation_effects_model, newdata = plot_data)

# Plotting with ggplot2
ggplot(plot_data, aes(x = avg_time_on_sm, y = predicted_score, color = occ_status, group = occ_status)) +
  geom_line() +
  labs(title = "Interaction Effects",
    x = "Average Time on Social Media", y = "Predicted Score") +
  scale_color_manual(values = c("blue", "green", "red"), name = "Occupation Status") +
  theme_minimal()
```
#### Checking which model is better
```{r}
glance(main_effects_model)$adj.r.squared
glance(intercation_effects_model)$adj.r.squared
```
#### Conclusion: By assuming that relationship between average time on social media and mental health score varies by the occupational groups we get a slightly more accurate model. Therefore, we will anaylze each occupational group seperatly.


# Univesity students
```{r}
# Filter data for university students
university_data <- data %>%
  filter(occ_status == "University Student")

# Fit the linear regression model
model_university <- lm(overall_mental_health_score ~ avg_time_on_sm + age + rel_status + gender + private_affiliated + not_affiliated + school_affiliated + 
              company_affiliated + government_affiliated + 
              uses_facebook + uses_twitter + uses_instagram + uses_youtube + 
              uses_discord + uses_reddit + uses_pinterest + uses_tiktok + 
              uses_snapchat, data = university_data)

# Summarize the model
summary(model_university)
```

```{r}
# For university students, explore the relationship between average time on social media and overall mental health score between users and non-users fo tiktok

# Fit the linear regression model
model_university <- lm(overall_mental_health_score ~ avg_time_on_sm+ uses_tiktok + age, data = university_data)

# Summarize the model
summary(model_university)
```

```{r}
# Predict on test data
test_university_data <- data %>%
  filter(occ_status == "University Student")

test_university_data <- test_university_data %>%
  mutate(predictions = predict(model_university, newdata = .))

# Calculate performance metrics on test data
rmse_university <- yardstick::rmse_vec(test_university_data$overall_mental_health_score, test_university_data$predictions)
r_squared_university <- yardstick::rsq_vec(test_university_data$overall_mental_health_score, test_university_data$predictions)

cat("University Students Model Performance:\n")
cat("RMSE:", rmse_university, "\n")
cat("R-squared:", r_squared_university, "\n")

# Get confidence intervals
conf_intervals <- confint(model_university)

# Combine results into a tidy data frame
model_tidy <- tidy(model_university) %>%
  mutate(
    conf.low = conf_intervals[, 1],
    conf.high = conf_intervals[, 2]
  )

# Calculate the Margin of Error
model_tidy <- model_tidy %>%
  mutate(margin_of_error = (conf.high - conf.low) / 2)

# Summarize and Display the Results
model_tidy %>%
  select(term, estimate, std.error, margin_of_error, conf.low, conf.high) %>%
  arrange(desc(abs(estimate))) %>%
  print()
```

```{r}
# Check linearity
ht_wt_fit_aug <- augment(model_university)

# Residuals vs Fitted plot for the extended model (Check homoscedasticity)
ggplot(model_university, aes(.fitted, .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted Values for University Students",
       x = "Fitted Values",
       y = "Residuals")

# Normal Q-Q plot for the extended model (Check normality)
qqnorm(lm_model_extended$residuals)
qqline(lm_model_extended$residuals, col = "red")
title(main = "\nfor University Students", line = 0.8)

```

```{r}
# Results plot
# Filter data for the relevant occupation statuses
filtered_data <- data %>%
  filter(occ_status %in% c("University Student"))

# Convert 'uses_tiktok' to a factor
filtered_data <- filtered_data %>%
  mutate(uses_tiktok = as.factor(uses_tiktok))

# Reshape the data to long format for illness scores
long_data_illness <- filtered_data %>%
  select(occ_status, uses_tiktok, adhd_score, depression_score, self_esteem_score, anxiety_score) %>%
  pivot_longer(
    cols = c(adhd_score, depression_score, self_esteem_score, anxiety_score),
    names_to = "illness",
    values_to = "score"
  )

# Create the plot with facet_grid for horizontal arrangement
ggplot(long_data_illness, aes(x = illness, y = score, fill = uses_tiktok)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  facet_grid(occ_status ~ ., scales = "free_x") +
  labs(
    title = "Illness Scores by TikTok Usage and Occupation Status",
    x = "Illness",
    y = "Score",
    fill = "Uses TikTok"
  ) +
    scale_x_discrete(labels = c(
    adhd_score = "ADHD Score",
    depression_score = "Depression Score",
    self_esteem_score = "Self-esteem Score",
    anxiety_score = "Anxiety Score"
  )) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    plot.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    strip.text = element_text(size = 12, face = "bold")
  )
```

# Salaried Workers
```{r}
### Model for salaried workers
# Filter data for salaried workers
workers_data <- data %>%
  filter(occ_status == "Salaried Worker")

# Fit the linear regression model
model_workers <- lm(overall_mental_health_score ~ avg_time_on_sm + age + rel_status + gender + private_affiliated + not_affiliated + school_affiliated + 
              company_affiliated + government_affiliated + 
              uses_facebook + uses_twitter + uses_instagram + uses_youtube + 
              uses_discord + uses_reddit + uses_pinterest + uses_tiktok + 
              uses_snapchat, data = workers_data)

# Summarize the model
summary(model_workers)
```

```{r}
# Fit the linear regression model
model_worker <- lm(overall_mental_health_score ~ avg_time_on_sm  + uses_instagram + uses_pinterest + school_affiliated + uses_facebook, data = worker_data)

# Summarize the model
summary(model_worker)
```

```{r}
# Predict on test data
test_worker_data <- data %>%
  filter(occ_status == "Salaried Worker")

# Predict on test data for Salaried Workers
test_worker_data <- test_worker_data %>%
  mutate(predictions = predict(model_worker, newdata = .))

# Calculate performance metrics on test data
rmse_worker <- yardstick::rmse_vec(test_worker_data$overall_mental_health_score, test_worker_data$predictions)
r_squared_worker <- yardstick::rsq_vec(test_worker_data$overall_mental_health_score, test_worker_data$predictions)

cat("Salaried Workers Model Performance:\n")
cat("RMSE:", rmse_worker, "\n")
cat("R-squared:", r_squared_worker, "\n")

# Get confidence intervals
conf_intervals <- confint(model_worker)

# Combine results into a tidy data frame
model_tidy <- tidy(model_worker) %>%
  mutate(
    conf.low = conf_intervals[, 1],
    conf.high = conf_intervals[, 2]
  )

# Calculate the Margin of Error
model_tidy <- model_tidy %>%
  mutate(margin_of_error = (conf.high - conf.low) / 2)

# Summarize and Display the Results
model_tidy %>%
  select(term, estimate, std.error, margin_of_error, conf.low, conf.high) %>%
  arrange(desc(abs(estimate))) %>%
  print()
```

```{r}
# Check linearity
ht_wt_fit_aug <- augment(model_work)

# Residuals vs Fitted plot for the extended model (Check homoscedasticity)
ggplot(model_work, aes(.fitted, .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted Values for Salaried Workers",
       x = "Fitted Values",
       y = "Residuals")

# Normal Q-Q plot for the extended model (Check normality)
qqnorm(lm_model_extended$residuals)
qqline(lm_model_extended$residuals, col = "red")
title(main = "\nfor Salaried Workers", line = 0.8)

```

```{r}
# Filter data for the relevant occupation statuses
filtered_data <- data %>%
  filter(occ_status %in% c("Salaried Worker"))

# Convert 'avg_time_on_sm' to a factor
filtered_data <- filtered_data %>%
  mutate(avg_time_on_sm = as.factor(avg_time_on_sm))

# Reshape the data to long format for illness scores
long_data_illness <- filtered_data %>%
  select(occ_status, avg_time_on_sm, adhd_score, depression_score, self_esteem_score, anxiety_score) %>%
  pivot_longer(
    cols = c(adhd_score, depression_score, self_esteem_score, anxiety_score),
    names_to = "illness",
    values_to = "score"
  )

# Create the plot with facet_wrap for horizontal arrangement
ggplot(long_data_illness, aes(x = avg_time_on_sm, y = score, fill = avg_time_on_sm)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  facet_wrap(~ illness, scales = "free_y", labeller = as_labeller(illness_labels)) +
  labs(
    title = "Mental Health Scores by Average Time on Social Media for Salaried Workers",
    x = "Average Time on Social Media",
    y = "Score",
    fill = "Average Time on Social Media"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    plot.title = element_text(size = 13.8, face = "bold", margin = margin(t = 10, b = 10), hjust = 0.75),  # Adjust the title margin
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    strip.text = element_text(size = 12, face = "bold")  )
```

# School students
```{r}
### Model for school student
# Filter data for university students
school_data <- train_data %>%
  filter(occ_status == "School Student")

# Fit the linear regression model
model_school <- lm(overall_mental_health_score ~ avg_time_on_sm + age + rel_status + gender + private_affiliated + not_affiliated + school_affiliated + 
              company_affiliated + government_affiliated + 
              uses_facebook + uses_twitter + uses_instagram + uses_youtube + 
              uses_discord + uses_reddit + uses_pinterest + uses_tiktok + 
              uses_snapchat, data = school_data)

# Summarize the model
summary(model_school)
```

```{r}
# Fit the linear regression model
model_school <- lm(overall_mental_health_score ~ company_affiliated + uses_facebook + avg_time_on_sm,  data = school_data)

# Summarize the model
summary(model_school)
```

```{r}
# Predict on test data
test_school_data <- data %>%
  filter(occ_status == "School Student")

# Predict on test data for Salaried Workers
test_school_data <- test_school_data %>%
  mutate(predictions = predict(model_worker, newdata = .))

# Calculate performance metrics on test data
rmse_school <- yardstick::rmse_vec(test_school_data$overall_mental_health_score, test_school_data$predictions)
r_squared_school <- yardstick::rsq_vec(test_school_data$overall_mental_health_score, test_school_data$predictions)

cat("School Students Model Performance:\n")
cat("RMSE:", rmse_school, "\n")
cat("R-squared:", r_squared_school, "\n")

# Get confidence intervals
conf_intervals <- confint(model_school)

# Combine results into a tidy data frame
model_tidy <- tidy(model_school) %>%
  mutate(
    conf.low = conf_intervals[, 1],
    conf.high = conf_intervals[, 2]
  )

# Calculate the Margin of Error
model_tidy <- model_tidy %>%
  mutate(margin_of_error = (conf.high - conf.low) / 2)

# Summarize and Display the Results
model_tidy %>%
  select(term, estimate, std.error, margin_of_error, conf.low, conf.high) %>%
  arrange(desc(abs(estimate))) %>%
  print()
```

```{r}
# Check linearity
ht_wt_fit_aug <- augment(model_school)

# Residuals vs Fitted plot for the extended model (Check homoscedasticity)
ggplot(model_school, aes(.fitted, .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted Values for School Students",
       x = "Fitted Values",
       y = "Residuals")

# Normal Q-Q plot for the extended model (Check normality)
qqnorm(lm_model_extended$residuals)
qqline(lm_model_extended$residuals, col = "red")
title(main = "\nfor School Students", line = 0.8)
```

```{r}
# Filter data for school students
school_data <- data %>%
  filter(occ_status == "School Student")

# Convert 'uses_facebook' to a factor
school_data <- school_data %>%
  mutate(uses_facebook = as.factor(uses_facebook))

# Reshape the data to long format for illness scores
long_data_illness <- school_data %>%
  select(occ_status, uses_facebook, adhd_score, depression_score, self_esteem_score, anxiety_score) %>%
  pivot_longer(
    cols = c(adhd_score, depression_score, self_esteem_score, anxiety_score),
    names_to = "illness",
    values_to = "score"
  )

# Create the plot with facet_grid for horizontal arrangement
ggplot(long_data_illness, aes(x = illness, y = score, fill = uses_facebook)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  facet_grid(occ_status ~ ., scales = "free_x") +
  labs(
    title = "Illness Scores by Facebook Usage and Occupation Status",
    x = "Illness",
    y = "Score",
    fill = "Uses Facebook"
  ) +
    scale_x_discrete(labels = c(
    adhd_score = "ADHD Score",
    depression_score = "Depression Score",
    self_esteem_score = "Self-esteem Score",
    anxiety_score = "Anxiety Score"
  )) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    plot.title = element_text(size = 14, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    strip.text = element_text(size = 12, face = "bold")
  )
```